\emph{Light Exposure Behavior Assessment (LEBA)}: Develop of a novel instrument to capture light exposure-related behaviours

### 2021-09-04 manuscript.Rmd

```{r,  out.width = '100%', out.height= '100%', fig.align = 'center', fig.cap='Development and psychometric properties of LEBA '}
knitr::include_graphics('Flowchart1.png')
```

@@

We used R (version 4.1.0), including several R packages, for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used a polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (pa) a factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by the minimum residuals extraction method as well. We used a combination factor identification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers.  Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018;

@@

```{r facid, out.height= "50%",  fig.align='center', fig.cap='Factor Identification (A) Parallel analysis (B) Scree Plot, (C) Hull method', fig.show="hold",fig.ncol = 2, out.width="50%",  warning=FALSE}
knitr::include_graphics('parallel.png')
scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)
knitr::include_graphics('HUll method.jpeg')

@@

Three rounds of EFA starting with all 48 items were conducted and problematic items were gradually discarded (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We confirmed this five-factor latent structure using varimax rotation with a minimum residual extraction method (see the supplementary). Table\@ref(tab:TabEFA5) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Fig\@ref(fig:Residuals) showed `r apa(sumR,2, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018a]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution.

In the five-factor solution, the first factor contained three items and explained `r apa(var1,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F1.alpha,2,F)`). All the items in this factor stemmed from the individual's preference to use blue light filters in different light environments. The second factor contained six items and explained `r apa(var2,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F2.alpha,2,F)`). Items under this factor commonly investigate an individual's hours spent outdoor. The third factor contained five items and explained `r apa(var3,2,F)`% of the total variance. Items under this factor dealt with the specific behaviors pertaining to sleep. However, the internal consistency reliability coefficient was not satisfactory ($\alpha$ = `r apa(F3.alpha,2,F)`). The fourth factor contained five items and explained `r apa(var4,2,F)`% of the total variance with an  internal consistency coefficient, $\alpha$ = `r apa(F4.alpha,2,F)`. These five items stemmed from the behavior related to an individual's cellphone usage during the sleep-wakeup time. Lastly, the fifth factor contained six items and explained `r apa(var5,2,F)`% of the total variance.This factor tried to measure an individual's behavior lead by the awareness of light's influence on health. However, this factor showed unsatisfactory internal consistency reliability ($\alpha$ = `r apa(F5.alpha,2,F)`). It is essential to attain a balance between psychometric properties and the interpretability of the common themes when exploring the latent structure. As all of the emerged factors are highly interpretable, regardless of the apparent low reliability of the two factors,  we retain the five-factor solution with 23 items for our confirmatory factor analysis (CFA). Two items showed negative factor-loading (items 44 and 21). Upon inspection, it was understood that these items are negatively correlated to the common theme, and thus in the CFA analysis, we reversed the response code for these two items.

@@

LB.model.Cor.1 <- "F1 =~ item16 + item36 + item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a1*item32 +item35 + Ritem37 + item38+ a2*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1+ Ritem26

a1==a2"

@@

## Summary of Model 
summary(fit.Cor.1, fit.measures =TRUE,standardized = TRUE)

## Selected Fit measures 
fit.1 <- fitmeasures (fit.Cor.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", 
                       "srmr"))

semTools::reliability(fit.Cor.1)

@@

```{r CFA23withMI, include=FALSE}
LB.model.Cor.2 <- "F1 =~ item16 + item36 + item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a1*item32 +item35 + item38+ a2*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1
                       a1==a2 
                item30 ~~  item41"

#Ritem26 & 37 are removed

@@

reliability <- semTools::reliability(fit.Cor.2, return.total = T)
as.table(reliability)
## Summary of Model 
summary(fit.Cor.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.2 <- fitmeasures (fit.Cor.2,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", 
                       "srmr"))

@@

@@

>>>>

```{r FlowchartFig,  out.width = '100%', out.height= '100%', fig.align = 'center', fig.cap='Development and psychometric properties of LEBA '}
knitr::include_graphics('Flowchart1.png')
```

@@

We used R (version 4.1.0), including several R packages, for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used a polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (pa), a factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by the minimum residuals extraction method as well. We used a combination factor identification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers.  Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018;

@@

```{r factor_identification_fig, out.height= "50%",  fig.align='center', fig.cap='Factor Identification (A) Parallel analysis (B) Scree Plot, (C) Hull method', fig.show="hold",fig.ncol = 2, out.width="50%",  warning=FALSE}
knitr::include_graphics('parallel.png')
scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)
knitr::include_graphics('HUll method.jpeg')

@@

Three rounds of EFA with all 48 items were conducted, and problematic items were gradually discarded (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We confirmed this five-factor latent structure using varimax rotation with a minimum residual extraction method (see the supplementary). Table\@ref(tab:TabEFA5) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Fig\@ref(fig:Residuals) showed `r apa(sumR,2, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018a]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution.

In the five-factor solution, the first factor contained three items and explained `r apa(var1,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F1.alpha,2,F)`). All the items in this factor stemmed from the individual's preference to use blue light filters in different light environments. The second factor contained six items and explained `r apa(var2,2,F)`% of the total variance with a satisfactory internal reliability coefficient ($\alpha$ = `r apa(F2.alpha,2,F)`). Items under this factor commonly investigate an individual's hours spent outdoor. The third factor contained five items and explained `r apa(var3,2,F)`% of the total variance. Items under this factor dealt with the specific behaviors pertaining to sleep. However, the internal consistency reliability coefficient was not satisfactory ($\alpha$ = `r apa(F3.alpha,2,F)`). The fourth factor contained five items and explained `r apa(var4,2,F)`% of the total variance with an  internal consistency coefficient, $\alpha$ = `r apa(F4.alpha,2,F)`. These five items stemmed from the behavior related to an individual's cellphone usage during the sleep-wakeup time. Lastly, the fifth factor contained six items and explained `r apa(var5,2,F)`% of the total variance. This factor tried to measure an individual's behavior lead by the awareness of light's influence on health. However, this factor showed unsatisfactory internal consistency reliability ($\alpha$ = `r apa(F5.alpha,2,F)`). It is essential to attain a balance between psychometric properties and the interpretability of the common themes when exploring the latent structure. As all of the emerged factors are highly interpretable, regardless of the apparent low reliability of the two factors,  we retain the five-factor solution with 23 items for our confirmatory factor analysis (CFA). Two items showed negative factor-loading (items 44 and 21). Upon inspection, it was understood that these items are negatively correlated to the common theme, and thus in the CFA analysis, we reversed the response code for these two items.

@@

LB.model.Cor.1 <- "F1 =~ a1*item16 + item36 + a2*item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a3*item32 + item35 + item38+ a4*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1 
                       a1==a2
                       a3==a4"

@@

## Summary of Model 
cfa.sum.1 <- summary(fit.Cor.1, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.1 <- fitmeasures (fit.Cor.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

reliability1 <- semTools::reliability(fit.Cor.1)

@@

```{r CFA23withMI, include=FALSE}
LB.model.Cor.2 <- "F1 =~ a1*item16 + item36 + a2*item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a3*item32 + item35 + item38+ a4*item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1 
                       a1==a2
                       a3==a4
                item30 ~~  item41
                       "

#item26 & 37 are removed

@@

reliability2 <- semTools::reliability(fit.Cor.2, return.total = T)
## Summary of Model 
cfa.sum.2 <- summary(fit.Cor.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.2 <- fitmeasures (fit.Cor.2,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

@@

We conducted a categorical confirmatory factor analysis with robust weighted least square (WLSMV) estimator as our response data was in ordinary nature [@desjardinsHandbookEducationalMeasurement2018a]. Commonly used Model fit suggestions of [@huCutoffCriteriaFit1999] focused on (i) the comparative fit index (CFI;) (ii) the Tucker Lewis index (TLI) (CFI/TLI, $goodfit \geq .95$, $acceptable fit\geq .90$) (ii) the root mean square error of approximation (RMSEA; close to .06 or below), (iii) the standardized root mean square (SRMR; close to .08 or below) to estimate the model fit. The chi-square test is also used to estimate the absolute model fit. Table\@ref(tab:tabCfa) summarizes the fit indices of our fitted model. The fitted model failed to attain an absolute fit estimated by the chi-square test. However, the chi-square test is sensitive to sample size and not recommended to be used as the sole index of absolute model fit [@RN1231]. The other absolute fit index we obtained in our analysis was SRMR which does not work well with categorical data [@RN1272]. Subsequently, we judged the model fit based on the incremental fit indices: CFI, TLI and parsimony fit index-RMSEA. Our fitted model attained acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`); RMSEA = `r apa(fit.1[7],2,F)`[`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI]) with two imposed equity constrain on item pairs 32-33 and 19-17 . However SRMR value was higher than the guideline (SRMR = `r apa(fit.1[10],2,F)`).  Further by allowing one pair of items (30-41) to covary their error variance and discarding two item (item 37 & 26) for very low r-square value, our model attained best fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI])   and SRMR value (SRMR = `r apa(fit.2[10],2,F)`) was also close to the suggestions of @huCutoffCriteriaFit1999.

```{r tabCfa, results='asis'}
CFa.matrix <- matrix("NA", nrow =2, ncol = 9)
colnames(CFa.matrix) = c("Model", "Chi-Squre", "df", "CFI", "TLI", "RMSEA", "RMSEA 90% Lower CI", "RMSEA 90% Upper CI", "SRMR")
CFa.matrix[1,] = c("Five factor model", 
                   apa(cfa.sum.1$FIT[6],2,T), 
                   apa(cfa.sum.1$FIT[7],2,T), 
                   apa(cfa.sum.1$FIT[19],2,F), 
                   apa(cfa.sum.1$FIT[20],2,T),
                   apa(cfa.sum.1$FIT[27],2,T),
                   apa(cfa.sum.1$FIT[28],2,T),
                   apa(cfa.sum.1$FIT[29],2,T),
                   apa(cfa.sum.1$FIT[35],2,T))

CFa.matrix[2,] = c("Five factor model", 
                   apa(cfa.sum.2$FIT[6],2,T), 
                   apa(cfa.sum.2$FIT[7],2,T), 
                   apa(cfa.sum.2$FIT[19],2,F), 
                   apa(cfa.sum.2$FIT[20],2,T),
                   apa(cfa.sum.2$FIT[27],2,T),
                   apa(cfa.sum.2$FIT[28],2,T),
                   apa(cfa.sum.2$FIT[29],2,T),
                   apa(cfa.sum.2$FIT[35],2,T))

apa_table(CFa.matrix, caption = "Fit indices of CFA")

@@

```{r InvarianceAnalysis_data, include=FALSE}
invariance.data <- cbind(CFA.data, data$slypos_demographics_language.factor[429:690])
colnames(invariance.data)[52] <- "English_Speaking"

invariance.data %>% 
  group_by(English_Speaking) %>% 
  count()
```

```{r Invarianceshortcut}
#library("semTools")
invariance.model <- measurementInvariance(model = LB.model.Cor.2,
                                          data = invariance.data,
                                          group = "English_Speaking",
                                          ordered =   names(CFA.data),
                                          estimator = "WLSMV",
                                          strict = T)
invariance.model$fit.configural
```

```{r Invariance detail, include=FALSE}
# Configural invariance ####

LB.model.Cor.3 <- "F1 =~ a1*item16 + a2*item36 + item17
                 F2 =~ item11 + item10 + item12+ item7+ Ritem8+ item9
                 F3 =~  item27+ item3+ item40 + item30 + item41
                 F4 =~ a4*item32 +a3*item35 + item38+ item33 
                 F5 =~ item46+ item45+ item25+ item4+ item1 
                       a1==a2
                       a4==a3
                item30 ~~  item41
F4~~1*F4"

configural <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV")

summary(configural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (configural,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

weak <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV", 
            group.equal = "loadings")

summary(weak, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (weak,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

First.comp <- compareFit (configural, weak)
summary(First.comp)

strong <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV", 
              group.equal = c("loadings", "intercepts"))
summary(strong, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

Second.comp <- compareFit  (weak, strong)
summary(Second.comp)

strict <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
              group.equal = c("loadings", "intercepts", "residuals"))

summary(strict, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (strict,c("gfi", "agfi", "nfi","rfi", "cfi.robust","tli.robust","rmsea", "srmr","aic"))

Third.comp <- compareFit(strict,strong)
summary(Third.comp)

structural <-cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
                 group.equal = c("loadings", "intercepts", "residuals", "means","lv.variances","lv.covariances"))

summary(strucrural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (structural,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

fourth.comp <- compareFit  (structural,strict)
summary(fourth.comp)
comfit.par <- compareFit(configural, weak, strong, strict, structural)
summary (comfit.par)

#library("semTable")
models <-  list("Configural" = configural, 
                "Metric" = weak, 
                "Scalar" = strong, 
                "Residual" = strict, 
                "Structural"=structural)

Invariance.table <- compareLavaan(models,
              nesting = "Configural > Metric > Scalar > Residual  > Structural", 
              fitmeas = c("chisq", "df", "pvalue", "cfi.scaled","tli.scaled","rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr" ),
              scaled = TRUE,
              chidif = TRUE, digits = 2)

```{r Invariance Tab}
apa_table(Invariance.table)
```
